### 2020-03-05 paper/si.tex

\documentclass[%
 reprint,
 superscriptaddress,
 amsmath,amssymb,
 onecolumn 
 ]{revtex4-1}

\usepackage{graphicx}
\usepackage{dcolumn}
\usepackage{bm}
\usepackage[separate-uncertainty=true]{siunitx}
\usepackage{mhchem}

\newcommand{\beginsupplement}{%
        \setcounter{table}{0}
        \renewcommand{\thetable}{S\arabic{table}}%
        \setcounter{figure}{0}
        \renewcommand{\thefigure}{S\arabic{figure}}%
     }

\begin{document}

\preprint{APS/123-QED}

\title{Supplementary information for ``A general approach to maximise information density in neutron reflectometry analysis''}

\author{Andrew R. McCluskey}
  \email{andrew.mccluskey@diamond.ac.uk}
  \email{a.r.mccluskey@bath.ac.uk}
  \affiliation{Diamond Light Source, Rutherford Appleton Laboratory, Harwell Science and Innovation Campus, Didcot, OX11 0DE, UK}
  \affiliation{Department of Chemistry, University of Bath, Claverton Down, Bath BA2 7AY, UK}

\author{Joshaniel F. K. Cooper}
  \affiliation{ISIS Pulsed Neutron and Muon Source, Science and Technology Facilities Council, Rutherford Appleton Laboratory, Harwell Science and Innovation Campus, Didcot, Oxfordshire, OX11 0QX, UK}

\author{Tom Arnold}
  \affiliation{European Spallation Source, SE-211 00 Lund, Sweden}
  \affiliation{Department of Chemistry, University of Bath, Claverton Down, Bath BA2 7AY, UK}
  \affiliation{ISIS Pulsed Neutron and Muon Source, Science and Technology Facilities Council, Rutherford Appleton Laboratory, Harwell Science and Innovation Campus, Didcot, Oxfordshire, OX11 0QX, UK}

\author{Tim Snow}
  \affiliation{Diamond Light Source, Rutherford Appleton Laboratory, Harwell Science and Innovation Campus, Didcot, OX11 0DE, UK}
  \affiliation{School of Chemistry, University of Bristol, Bristol, BS8 1TS, UK}

\date{\today}

\maketitle
\beginsupplement

\section{Evidence}

The Bayesian $\ln\{p(\mathbf{D}|H)\}$ for each of the possible permutations of free parameters investigated in this work is given in Table~\ref{tab:evidence}.

%
\begin{table*}[h]
\caption{\label{tab:evidence} The values for evidence for each of the possible permutations of the 6 free parameters.}
\begin{ruledtabular}
\begin{tabular}{lrlr}
Free parameters & $p(\bm{D}|H)$ & Free parameters & $p(\bm{D}|H)$ \\
\colrule
\input{ev_table.txt}\unskip
\end{tabular}
\end{ruledtabular}
\end{table*}
%

\end{document}